---
title: "GPDI517 Data Visualization Worksheet"
author: "Tugba Ozturk"
date: "March 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Viz with base-R:

We will use `iris` data frame. Before diving into the `ggplot` world, let's look at the data set and plot a simple graph using `plot()` function:

```{r df}
#head(iris)
df <- iris
colnames(df) <- c("l","w","pl","pw","spe")
```

Let's load `dplyr` and create a new column by multiplying the `r colnames(df)[3]` and `r colnames(df)[4]` columns. Make sure to use `head()` function to check `df` data frame and its new column.

```{r dplyrpractice}
 #loads the dplyr package
 #creates a new column named plpw
 #shows the first 10 lines of the updated data frame
```

Now, let's try out the `plot()` function. 

```{r plot}
plot(df$l, df$plpw, type="l") #change the code so that it generates a scatter plot!
```

Add x and y labels to the plot, change the color of the data points as well as the symbol. Make sure that the range of its y axis is [0,25].

Let's look at the following code chunk and add comments to explain each line:

```{r plotpractice}
label <- ifelse(df$plpw>10,"salmon","skyblue3")
with(df, plot(l, plpw, type="p", col=label))
abline(h=10, lty=2, lwd=1.2)
legend("topleft", c("expected", "rare"), fill=c("skyblue3","salmon"), bty="n")
```

## ggplot:


Now, install and load the `ggplot` package and plot a graph with `ggplot`. Be careful to use a data frame; `ggplot` works best with data frames!

```{r plotpractice2}
#install.packages("ggplot2")
library("ggplot2")
head(mtcars, 15) #let's remember the data set!
myplot <- ggplot(mtcars, aes(x=mpg, y=disp, size=hp, col=as.factor(cyl), shape=as.factor(gear))) + geom_point() + 
  labs(x="miles per gallon", y="displacement", size="horsepower", 
       col="# of cylinders", shape="# of gears")

print(myplot)
```

You can add `theme_bw()` to change the theme of your plot as black and white. Check [this website](http://sape.inf.usi.ch/quick-reference/ggplot2/themes) to learn about other `ggplot` themes.

Similarly, fix the following chunk carefully until it generates a graph combining *wt* vs *disp* and *mpg* vs *disp*:

```{r plotpractice3}
myploy <- ggplot() + 
  geom_line(data = mtcars, aes(, , color = "wt"),  lty=1) + 
  geom_line(data = mtcars, aes(, , color = "mpg"),  lty=2) + 
  ylab("wt or mpg") + xlab("disp") +
  scale_color_manual(name = "My legend:", 
                     values = c("wt" = "", "mpg" = "")) +
  scale_linetype_manual(values=c("solid","dashed")) +
  theme(axis.text.x=element_text(size=rel(1.2), angle=90),
        axis.text.y=element_text(size=rel(1.2)), panel.grid.major = , 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        legend.key=  ) 
print()
```


Now, let's use `qplot()` function from the `ggplot` package to plot a histogram. Check the following R code and modify the code 

- to set its color to `seagreen3`;
- to make the bars transparent;
- to set its theme to black and white;
- to save the plot as a png file.

```{r plotpractice4}

P <- qplot(mtcars$drat, geom="histogram", binwidth= 0.2, fill=I("blue"), col=I("blue"), bins=20, alpha=) + theme(text = element_text(size=12), legend.position = "none") + xlab("D") + ylab("Count") 

```

One last example with `ggplot` which requires us to read two data frames: *wildtype.dat* and *mutant.dat*.

```{r ggplotlastexample1}
wildtype <- read.table(file="wildtype.dat")
mutant <- read.table("mutant.dat")
```

Check the column names of these data frames. Then, use the following code as a template and follow the given steps:

- combine wildtype and mutant data frames and name the data frame as `all`;
- use `ggplot()` function to plot `psi` vs `phi` graph and color it by the `chain` column;
- make sure to generate separate plots by the combination of `prot` and `ind` columns (modify inside of the `facet_wrap()` function!).


```{r ggplotlastexample2}
     #combine the data frames
print(ggplot(data=all, aes(x=, y=, col=))+geom_point(size=0.5) + 
        facet_wrap(~paste(  ,  ,  sep="-"), ncol=2) + 
        xlab(expression(paste(phi, " (",degree,")")))+ 
        ylab(expression(paste(psi," (", degree, ")")))+
        theme_bw() +
        theme_grey(base_size = 18))
```


## Plotly:
Let's load the `plotly` package to generate an interactive histogram graph of the price variable from the `diamonds` data set:

```{r plotpractice5}

library("plotly")
set.seed(80)
d <- diamonds[sample(nrow(diamonds), 500), ]
h <- plot_ly(d, x=~price, type="", color="")
print(h)
layout(h, title="Diamond Prices", yaxis=list(title="Count"), xaxis=list(title="Price"))
```

Generate the plot using the following code chunk:

```{r plotpractice6}
cvsp <- plot_ly(d, x=~price, y=~carat, type="scatter", mode="markers")
```

Then add the following arguments:

- color= ~cut
- size=~depth
- text = ~paste("Clarity: ", clarity)

Carefully add the comments to explain the changes in your plot. 