---
title: "GPDI517: Introduction to data visualization"
author: "by Tugba Ozturk"
date: "April 4, 2019"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Viz with base-R:

Before diving into the `ggplot` world, let's start with the `iris` data frame with base-R and plot a simple graph using `plot()` function:

```{r df}
#head(iris)
df <- iris
colnames(df) <- c("l","w","pl","pw","spe")
```

Let's load `dplyr` and create a new column by multiplying the `r colnames(df)[3]` and `r colnames(df)[4]` columns. Make sure to use `head()` function to check `df` data frame and its new column.

```{r dplyrpractice}
library("dplyr") #loads the dplyr package
df <- df %>% mutate(plpw=pl*pw) #creates a new column named plpw
head(df,10) #shows the first 10 lines of the updated data frame
```

Now, let's try out the `plot()` function. Examine the given code and modify it to add x and y labels to the plot, change the color of the data points as well as the symbol. Make sure that the range of its y axis is [0,25]. Hint: `expression(paste())` is a useful combination for the labels!!!

```{r plot}
plot(df$l, df$plpw, type="l") #change the code so that it generates a scatter plot!
#type="l" for line, "b" for both!


```


Let's look at the following code chunk and add comments to explain each line:

```{r plotpractice}
label <- ifelse(df$plpw>10,"salmon","skyblue3")
with(df, plot(l, plpw, type="p", col=label))
abline(h=10, lty=2, lwd=1.2)
legend("topleft", c("expected", "rare"), fill=c("skyblue3","salmon"), bty="n")
```

## ggplot:

Now, install and load the `ggplot` package and plot a graph with `qplot()` and `ggplot()` functions. Be careful to use a data frame; `ggplot` works best with data frames! Examine the examples given below:

```{r ggplotpractice1}
library("ggplot2")
#use color brewer as default discrete colors
scale_colour_discrete <- function(...) scale_color_brewer(palette="Set1", ...)
scale_fill_discrete <- function(...) scale_fill_brewer(palette="Set1", ...)
data('mtcars')
mtcars$gear <- factor(mtcars$gear,levels=c(3,4,5),
   labels=c("3gears","4gears","5gears")) 
mtcars$am <- factor(mtcars$am,levels=c(0,1),
     labels=c("Automatic","Manual"))
mtcars$cyl <- factor(mtcars$cyl,levels=c(4,6,8),
   labels=c("4cyl","6cyl","8cyl"))


qplot(mpg, data=mtcars, geom="density", fill=gear, alpha=I(.5), 
   main="Distribution of Gas Milage", xlab="Miles Per Gallon", 
   ylab="Density") 

qplot(hp, mpg, data=mtcars, shape=am, color=am,
   facets=gear~cyl, size=I(3),
   xlab="Horsepower", ylab="Miles per Gallon")
```


```{r plotpractice2}
#install.packages("ggplot2")
library("ggplot2")
head(mtcars, 15) #let's remember the data set!
myplot <- ggplot(mtcars, aes(x=mpg, y=disp, size=hp, col=as.factor(cyl), shape=as.factor(gear))) + geom_point() + 
  labs(x="miles per gallon", y="displacement", size="horsepower", 
       col="# of cylinders", shape="# of gears")

print(myplot)
```

You can add `theme_bw()` to change the theme of your plot as black and white. Check [this website](http://sape.inf.usi.ch/quick-reference/ggplot2/themes) to learn about other `ggplot` themes.

Fix the following chunk carefully until it generates a complete graph (having two line plots: wt vs disp and mpg vs disp):

```{r plotpractice3}
#ggplot() + 
#  geom_line(data = mtcars, aes(, , color = "wt"),  lty=) + 
#  geom_line(data = mtcars, aes(, , color = "mpg"),  lty=) + 
#  ylab("wt or mpg") + xlab("disp") +
#  scale_color_manual(name = "My legend:", 
#                     values = c("wt" = "blue", "mpg" = "orange")) +
#  scale_linetype_manual(values=c("solid","dashed")) +
#  theme(axis.text.x=element_text(size=rel(1.2), angle=180),   axis.text.y=element_text(size=rel(1.2)), panel.grid.major = element_blank(), panel.grid.minor = ,
#panel.background = , 
#axis.line = element_line(colour = "yellow"),
#legend.key=) 
  
```


Now, let's use `qplot()` function from the `ggplot` package to plot a histogram. Check the following R code and modify the code 

- to set its color to `seagreen3`;
- to make the bars transparent;
- to set its theme to black and white;
- to save the plot as a png file.

```{r plotpractice4}
#

#P <- qplot(mtcars$drat, geom="histogram", col=, binwidth= 0.2, bins=20, alpha=) + theme(text = element_text(size=12), legend.position = "none") + xlab("D") + ylab("Count")
#print(P)
#
```

One last example with `ggplot()` which requires us to read two data frames: *wildtype.dat* and *mutant.dat*.

```{r ggplotlastexample1}
wildtype <- read.table(file="wildtype.dat")
mutant <- read.table("mutant.dat")
```

Check the column names of these data frames. Then, use the following code as a template and follow the given steps:

- combine wildtype and mutant data frames and name the data frame as `all`;
- use `ggplot()` function to plot `psi` vs `phi` graph and color it by the `chain` column;
- make sure to generate separate plots by the combination of `prot` and `ind` columns (modify inside of the `facet_wrap()` function!).


```{r ggplotexample2}

#print(ggplot(data=all, aes(x=, y=, col=))+geom_point(size=0.5)+facet_wrap(~paste(,sep=""),ncol=2)+xlab(expression(paste(phi, " (",degree,")")))+ylab(expression(paste(psi," (", degree, ")")))+theme_bw() + theme_grey(base_size = 18)) #base_size increases the size of all text in the graph consistently!

```


Example three:
```{r ggplotexample3}


```


## Plotly:
Let's load the `plotly` package to generate an interactive histogram graph of the price variable from the `diamonds` data set:

```{r plotpractice5}

library("plotly")
set.seed(80)
d <- diamonds[sample(nrow(diamonds), 500), ]
h <- plot_ly(d, x=~price, type="histogram", color="salmon")
print(h)
layout(h, title="Diamond Prices", yaxis=list(title="Count"), xaxis=list(title="Price"))
```

Generate the plot using the following code chunk, then add the following arguments:

- color= ~cut
- size= ~z
- text = ~paste("Clarity: ", clarity, "<br>Price: ", price, "$")

Carefully add the comments to explain the changes in your plot. 

```{r , fig.width = 8, fig.height = 8}
#df <- diamonds[sample(1:nrow(diamonds), size = 2000),]
 
#plot_ly(df, x = ~x, y = ~price, mode = "markers") %>%
#layout(title = "Diamonds")
```

Visit [Plotly's webpage](https://plot.ly/r/line-and-scatter/) for more information.
